---
title: "DATA 624 Week 4"
author: "Vinicio Haro"
date: "September 25, 2019"
output: html_document
---

#Chapter 4, Forecasting Principles and Practice, HA
## problems 7.1 & 7.3


## 7.1 Consider the pigs series — the number of pigs slaughtered in Victoria each month

* a) Use the ses() function in R to find the optimal values of  alpha and l0 , and generate forecasts for the next four months.

```{r warning=FALSE, message=FALSE}
library(fpp2)
library(ggplot2)

#summary(pigs)

summary(ses(pigs,h=4))
```

From the output, we observe the alpha to be 0.2971 and l to be 10308.58

* b) Compute a 95% prediction interval for the first forecast using y±1.96 s where s is the standard deviation of the residuals. Compare your interval with the interval produced by R.

```{r warning=FALSE, message=FALSE}

s<-sd((ses(pigs, h=4))$residuals)
print(paste0("lower Confidence Interval: ", ses(pigs,h=4)$mean[1]-1.96*s))
print(paste0("Upper Confidence Interval: ", ses(pigs,h=4)$mean[1]+1.96*s))


```

Our confidence intervals are slightly different than the ones produced by r's output. They seem to be more narrow. 

## 7.3
Modify your function from the previous exercise to return the sum of squared errors rather than the forecast of the next observation. Then use the optim() function to find the optimal values of  
alpha and lo. Do you get the same values as the ses() function?

We clearly can't do this problem without doing 7.2 first which is...
Write your own function to implement simple exponential smoothing. The function should take arguments y (the time series), alpha (the smoothing parameter and level. It should return the forecast of the next observation in the series. Does it give the same forecast as ses()?

```{r warning=FALSE, message=FALSE}
my_ses <- function(y, alpha, l0)
  {
  y_hat <- l0
  for(index in 1:length(y))
    {
   y_hat <- alpha*y[index] + (1 - alpha)*y_hat 
    }
  cat("Forecast result by ses_optim function: ",
      as.character(y_hat),
      sep = "\n")
  }

```











